Universal relations after GW170817

### 2018-11-27 paper/bin-love-after-gw170817.tex
added comments

@@

\subsection{Spectral representations of neutron star equations of state}\label{sec:eos}
The structure of a NS and its tidal interactions in a binary system rely heavily on the underlying state function describing the relationship between pressure ($p$) and energy density ($\epsilon$), or equation of state, of nuclear matter.
Given that all currently proposed EoSs utilize certain approximations, one method to study a wide range of physically realizeable EoSs is to parameterize them in a model-independent way.
Spectral representations~\cite{Lindblom:2010bb,Lindblom:2012zi,Lindblom:2013kra,Lindblom:2018rfr,Abbott:2018exr} parameterize EoSs by performing spectral expansions on the adiabatic index $\Gamma(p)$\footnote{Another way of parameterizing EoSs is the piecewise polytropic formulation~\cite{Read2009,Lackey:2014fwa,Carney:2018sdv}.}:
\begin{equation}
\Gamma(x) = \exp{\sum_k\gamma_k x^k},
\end{equation}
where $x \equiv \log{p/p_0}$ for minimum pressure $p_0$.
The equation of state is then determined by an integration of the differential equation:
\begin{equation}
\frac{d \epsilon(p)}{dp}=\frac{\epsilon(p)+p}{p \Gamma(p)}.
\end{equation}9
Using this formalism, any valid EoS can be approximated through choice of 4 or more spectral coefficients $\gamma_k$, tabulated for several common EoSs in Table 1 of~\cite{Lindblom:2018rfr}.

In the current analysis, we consider only EoSs which have been constrained by the recent binary neutron star merger event, GW170817.
Important recent work~\cite{LIGO:posterior} uniformly sampled EoS parameters in order to derive a marginalized posterior on the pressure as a function of mass density, as seen in Fig. 2 of~\cite{LIGO:posterior}.
In our work, we similarly sample spectral coefficients within the ranges: $\gamma_0 \in \lbrack 0.2,2 \rbrack$, $\gamma_1 \in \lbrack -1.6,1.7 \rbrack$, $\gamma_2 \in \lbrack -0.6,0.6 \rbrack$, $\gamma_0 \in \lbrack -0.02,0.02 \rbrack$, and further restricting the adiabatic index to be $\Gamma \in \lbrack 0.6,4.5 \rbrack$, ensuring our parameterization exposes a wide range of viable EoSs~\cite{Lindblom:parameters}.
In addition, we impose similar constraints as was done in~\cite{LIGO:posterior}: (i) causality within 10\%, and (ii) EoS priors must support NS masses up to $1.97 \text{ M}_{\odot}$, consistent with astrophysical observations.
We then utilize the posterior defined by GW170817 by rejecting any EoS sample which does not lie within the 90\% credible posterior region, resulting in a sample of 35 EoSs shown in Fig.~\ref{fig:eos}.
Following~\cite{Read2009}, the parameterized high-density core EoSs generated above are matched to the low-density crust EoS Sly~\cite{Douchin:2001sv} at about half of the nuclear saturation density, $\rho_{\text{stitch}}=1.3 \times 10^{14} \text{ g/cm}^3$.
For comparison, we also include a second sample of randomly generated EoSs, without restricting to GW170817s posterior.
\begin{figure}
\begin{center} 
\includegraphics[width=\columnwidth]{EoSs.eps}
\end{center}
\caption{
Representative set of EoSs used in our analysis. 
These were populated by uniformly sampling the spectral coefficient parameter space, then restricting to the 90\% credible level shown in cyan~\cite{LIGO:posterior}.
In addition, the following constraints were also applied: (i) the EoS must have a causal structure within 10\%, and (ii) the EoS must support maximum NS mass of at least $1.97 \text{ M}_{\odot}$, consistent with astrophysical observations.
Finally, the high density core EoSs generated above were stitched to the low-density crust EoS of SLy~\cite{Douchin:2001sv} at $\rho_{\text{stitch}}=1.3 \times 10^{14} \text{ g/cm}^3$.
}
\label{fig:eos}
\end{figure}

@@

Fig.~\ref{fig:ILQ} shows the improved I-Love universal relation between dimensionless moment of inertia $\bar{I} \equiv I/M^3$ and dimensionless tidal deformability $\Lambda$ for both constrained and unconstrained sets of EoSs.
We observe that the maximum recorded EoS variation under the constrained set of EoSs is $0.43\%$, showing considerable improvement from the $0.88\%$ found in the case of unconstrained EoSs.
Further, both EoS samples show improvement from the original result of $\approx 1\%$ displayed in~\cite{Yagi:ILQ}.
\begin{figure*}
\begin{center} 
\includegraphics[width=\textwidth]{ILQ.eps}
\end{center}
\caption{
I-Love universal relations shown for the constrained EoSs (left) and unconstrained EoSs (right).
In these figures, the black dashed lines corresponds to the fit given by Eq.~\ref{eq:ILQfit}, while all other colors correspond to the 35 EoSs used in each case.
The maximal EoS variation from the fit for unconstrained EoSs is $0.88\%$, which is comparable to the $1\%$ found in~\cite{Yagi:ILQ}.
The constrained EoSs on the other hand, are universal to $0.43\%$, showing considerable improvement upon previous analyses.
}
\label{fig:ILQ}
\end{figure*}

@@

@@

@@

@@

>>>>

\definecolor{red(ncs)}{rgb}{0.77, 0.01, 0.2}
\newcommand{\ny}[1]{\textcolor{blue}{NY: #1} }

@@

\subsection{Spectral representations of neutron star equations of state}\label{sec:eos}
The structure of a NS and its tidal interactions in a binary system rely heavily on the underlying state function describing the relationship between pressure ($p$) and energy density ($\epsilon$), or equation of state, of nuclear matter.
Given that all currently proposed EoSs utilize certain approximations, one method to study a wide range of physically realizeable EoSs is to parameterize them in a model-independent way.
Spectral representations~\cite{Lindblom:2010bb,Lindblom:2012zi,Lindblom:2013kra,Lindblom:2018rfr,Abbott:2018exr} parameterize EoSs by performing spectral expansions on the adiabatic index $\Gamma(p)$\footnote{Another way of parameterizing EoSs is the piecewise polytropic formulation~\cite{Read2009,Lackey:2014fwa,Carney:2018sdv}.}:
\begin{equation}
\Gamma(x) = \exp{\sum_k\gamma_k x^k},
\end{equation}
where $x \equiv \log{p/p_0}$ for minimum pressure $p_0$.
The equation of state is then determined by an integration of the differential equation:
\begin{equation}
\frac{d \epsilon(p)}{dp}=\frac{\epsilon(p)+p}{p \Gamma(p)}.
\end{equation}
Using this formalism, any valid EoS can be approximated through choice of 4 or more spectral coefficients $\gamma_k$, tabulated for several common EoSs in Table 1 of~\cite{Lindblom:2018rfr}.

In the current analysis, we consider only EoSs which have been constrained by the recent binary neutron star merger event, GW170817.
Important recent work~\cite{LIGO:posterior} uniformly sampled EoS parameters in order to derive a marginalized posterior on the pressure as a function of mass density, as seen in Fig. 2 of~\cite{LIGO:posterior}.
In our work, we similarly sample spectral coefficients within the ranges: $\gamma_0 \in \lbrack 0.2,2 \rbrack$, $\gamma_1 \in \lbrack -1.6,1.7 \rbrack$, $\gamma_2 \in \lbrack -0.6,0.6 \rbrack$, $\gamma_0 \in \lbrack -0.02,0.02 \rbrack$, and further restricting the adiabatic index to be $\Gamma \in \lbrack 0.6,4.5 \rbrack$, ensuring our parameterization exposes a wide range of viable EoSs~\cite{Lindblom:parameters}.
In addition, we impose similar constraints as was done in~\cite{LIGO:posterior}: (i) causality within 10\%, and (ii) EoS priors must support NS masses up to $1.97 \text{ M}_{\odot}$, consistent with astrophysical observations.
We then utilize the posterior defined by GW170817 by rejecting any EoS sample which does not lie within the 90\% credible posterior region, resulting in a sample of 35 EoSs shown in Fig.~\ref{fig:eos}.
Following~\cite{Read2009}, the parameterized high-density core EoSs generated above are matched to the low-density crust EoS Sly~\cite{Douchin:2001sv} at about half of the nuclear saturation density, $\rho_{\text{stitch}}=1.3 \times 10^{14} \text{ g/cm}^3$.
For comparison, we also include a second sample of randomly generated EoSs, without restricting to GW170817s posterior. {\ny{It would be good to make this figure a double pane one, so that you can show the constrained set and the unconstrained set side-by-side.}}
\begin{figure}
\begin{center} 
\includegraphics[width=\columnwidth]{EoSs.eps}
\end{center}
\caption{
Representative set of EoSs used in our analysis. 
These were populated by uniformly sampling the spectral coefficient parameter space, then restricting to the 90\% credible level shown in cyan~\cite{LIGO:posterior}.
In addition, the following constraints were also applied: (i) the EoS must have a causal structure within 10\%, and (ii) the EoS must support maximum NS mass of at least $1.97 \text{ M}_{\odot}$, consistent with astrophysical observations.
Finally, the high density core EoSs generated above were stitched to the low-density crust EoS of SLy~\cite{Douchin:2001sv} at $\rho_{\text{stitch}}=1.3 \times 10^{14} \text{ g/cm}^3$.
}
\label{fig:eos}
\end{figure}

@@

{\ny{Let me play devil's advocate and explain things I could potentially be worried about.
\begin{itemize}
\item The constrained and the unconstrained data set of I-Love that you are fitting, do they have the same number of data points? If not, how do you know your fits and the comparisons are reliable to variation on the data set? You can study this by re-doing this analysis with the same amount of data in both sets of EoSs and then repeating the analysis with twice the amount of data to see how the results change.  
\item  The unconstrained set does not contain the same EoSs as the constrained set plus additional ones that are ruled out. This makes comparison of the sets a bit more difficult, because they are randomly sampled. How do you know your comparisons are robust to the number of EoSs you have randomly chosen without your probability distribution? You can study this by re-doing this analysis with twice the number of drawn EoSs to see how the results change. 
\item In the constrained set, there seems to be 3 EoSs that have the ``largest'' difference. What makes these EoSs special? Do they have anything in common?  Same question for the unconstrained data set, but there are 2 special ones this time. 
\end{itemize}
}}
Fig.~\ref{fig:ILQ} shows the improved I-Love universal relation between dimensionless moment of inertia $\bar{I} \equiv I/M^3$ and dimensionless tidal deformability $\Lambda$ for both constrained and unconstrained sets of EoSs.
We observe that the maximum recorded EoS variation under the constrained set of EoSs is $0.43\%$, showing considerable improvement from the $0.88\%$ found in the case of unconstrained EoSs.
Further, both EoS samples show improvement from the original result of $\approx 1\%$ displayed in~\cite{Yagi:ILQ}.

{\ny{This comment is related to the list above, but I'm worried about how robust these numbers quoted above to *2* sig figs are...}}

\begin{figure*}
\begin{center} 
\includegraphics[width=\textwidth]{ILQ.eps}
\end{center}
\caption{
I-Love universal relations shown for the constrained EoSs (left) and unconstrained EoSs (right).
In these figures, the black dashed lines corresponds to the fit given by Eq.~\ref{eq:ILQfit}, while all other colors correspond to the 35 EoSs used in each case.
The maximal EoS variation from the fit for unconstrained EoSs is $0.88\%$, which is comparable to the $1\%$ found in~\cite{Yagi:ILQ}.
The constrained EoSs on the other hand, are universal to $0.43\%$, showing considerable improvement upon previous analyses.
}
\label{fig:ILQ}
\end{figure*}

@@

{\ny{I think it would be worth studying whether there is a better fitting function that one can use for the constrained set. The functional form in Eq. (6) does not limit properly to the Newtonian result or to the relativistic result. But we do know how I-Love for example scales in the Newtonian limit, so could we e.g. peel off the asymptotic behavior analytically, and then correct this with a fitting function? This is similar to what Kent and I did in the past in a paper (in the binary love actually) and also similar to the way EoB works.}}

@@

{\ny{Why do we consider hybrid stars only in this subsection? Why not also include these in the I-Love studies?}}

@@

{\ny{I have very similar comments here to those I listed in the I-Love relation section.}}

@@

{\ny{The comparison below is a bit misleading because all you have shown is that the fitting coefficients for the standard stars do not work when dealing with mixed stars. I think it would be interesting to see how the binary love relations look when you include hybrid systems (something like Fig. 5 but for QS/QS and for QS/HS). Also I think it would be interesting to either (i) redo the fits including the QS/QS and QS/HS data and or (ii) produce different fits for those families, and then you can compare either of these to a wide set of systems to determine the validity of the fit.}}
